Practical Optional Types for Clojure

### 2015-02-23 formal-model.tex
formatting, remove/inline figures

\begin{figure*}
  \footnotesize
$$
\begin{altgrammar}
  \expd{}, \e{} &::=& \x{}
                      \alt \v{} 
                      \alt {\comb {\e{}} {\e{}}} 
                      \alt {\abs {\x{}} {\t{}} {\e{}}}
                      \alt {\ifexp {\e{}} {\e{}} {\e{}}}
                      %\alt {\trdiff{\doexp {\e{}} {\e{}}}}
                      \alt {\letexp {\x{}} {\e{}} {\e{}}}
                      %\alt {\errorvalv{}}
                &\mbox{Expressions} \\
  \v{} &::=&          \singletonmeta{}
                      \alt {\const{}}
                      \alt {\closure {\openv{}} {\abs {\x{}} {\t{}} {\e{}}}}
                &\mbox{Values} \\
                \constantssyntax{}\\
  \s{}, \t{}    &::=& \Top 
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt
                      {\ArrowOne {\x{}} {\t{}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                      \alt {\Value \singletonmeta{}} 
                      \alt \trdiff{\class{}}
                &\mbox{Types} \\
  \singletonallsyntax{}
                \\ \\
  \occurrencetypingsyntax{}\\
  \openvsyntax{}
  %\\
  %\classliteralallsyntax{}
\end{altgrammar}
$$
\caption{Syntax of Terms, Types, Propositions and Objects}
\label{main:figure:termsyntax}
\end{figure*}

@@

    {\TDo}
    %{\TClass}
    {\TIf}
    {\TLet}

@@

\begin{figure*}
  \footnotesize
  \begin{mathpar}
    {\BLocal}

    %{\BDo}

    {\BLet}

    \BVal{}

    \BIfTrue{}

    \BIfFalse{}

    \BAbs{}

    \BBetaClosure{}

    \BDelta{}
  \end{mathpar}
  \caption{Operational Semantics}
  \label{main:figure:standardopsem}
\end{figure*}

@@

\begin{figure*}
  \footnotesize
  \begin{mathpar}
    \RAbs{}

    \RNewElimRefl{}

    \RMethodElimRefl{}

    \RFieldElimRefl{}

    \RLet{}

    \RLetHint{}
  \end{mathpar}
\caption{Reflection Elimination (select rules, figure~\ref{appendix:figure:rewrite} for full rules)}
\label{main:figure:rewrite}
\end{figure*}

\begin{figure*}
  \footnotesize
  \begin{mathpar}
    \TALocal{}

    \TANil{}

    \TANewStatic{}

    \TALetHint{}

    \TALet{}
  \end{mathpar}
\caption{Type Hint Inference (select rules, figure~\ref{appendix:figure:hintinfer} for full rules)}
\label{main:figure:hintinfer}
\end{figure*}

@@

core.typed uses tools.analyzer to obtain a rich abstract syntax tree that includes resolved
reflection, however to tell the full story we include a rewrite relation 
in figure~\ref{main:figure:rewrite} to eliminate reflection when sufficient type hints are provided.
Figure~\ref{main:figure:hintinfer} shows the interesting cases in type hint inference.

@@

@@

{\footnotesize
\input{constant}
\input{update}
}

@@

>>>>

\begin{figure}
  \footnotesize
$$
\begin{array}{lrlr}
  \expd{}, \e{} &::=& \x{}
                      \alt \v{} 
                      \alt {\comb {\e{}} {\e{}}} 
                      \alt {\abs {\x{}} {\t{}} {\e{}}} &\mbox{Expressions} \\
                      &\alt& {\ifexp {\e{}} {\e{}} {\e{}}}
                      %\alt {\trdiff{\doexp {\e{}} {\e{}}}}
                      \alt {\letexp {\x{}} {\e{}} {\e{}}}\\
                      %\alt {\errorvalv{}}
  \v{} &::=&          \singletonmeta{}
                      \alt {\const{}}
                      \alt {\closure {\openv{}} {\abs {\x{}} {\t{}} {\e{}}}}
                &\mbox{Values} \\
                \constantssyntax{}\\
  \s{}, \t{}    &::=& \Top 
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt
                      {\ArrowOne {\x{}} {\t{}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                &\mbox{Types} \\
                      &\alt& {\Value \singletonmeta{}} 
                      \alt \trdiff{\class{}}\\
  \singletonallsyntax{}
                \\ \\
  \occurrencetypingsyntax{}\\
  \openvsyntax{}
  %\\
  %\classliteralallsyntax{}
\end{array}
$$
\caption{Syntax of Terms, Types, Propositions and Objects}
\label{main:figure:termsyntax}
\end{figure}

@@

    %{\TDo}
    %{\TClass}
    {\TIf}
    {\TLet}

@@

\begin{figure}
  \footnotesize
  \begin{mathpar}
%    {\BLocal}
%
%    %{\BDo}
%
%    {\BLet}
%
%    \BVal{}
%
    \BIfTrue{}

%    \BIfFalse{}
%
%    \BAbs{}
%
%    \BBetaClosure{}
%
%    \BDelta{}
  \end{mathpar}
  \caption{Operational Semantics}
  \label{main:figure:standardopsem}
\end{figure}

@@

\begin{figure}
  \footnotesize
  \begin{mathpar}
    %\RAbs{}
%
%    \RNewElimRefl{}
%
%    \RMethodElimRefl{}
%
    \RFieldElimRefl{}
%
%    \RLet{}
%
    \RLetHint{}
  \end{mathpar}
\caption{Reflection Elimination (select rules, figure~\ref{appendix:figure:rewrite} for full rules)}
\label{main:figure:rewrite}
\end{figure}

%\begin{figure}
%  \footnotesize
%  \begin{mathpar}
%%    \TALocal{}
%%
%%    \TANil{}
%%
%%    \TANewStatic{}
%%
%%    \TALetHint{}
%
%%    \TALet{}
%  \end{mathpar}
%\caption{Type Hint Inference (select rules, figure~\ref{appendix:figure:hintinfer} for full rules)}
%\label{main:figure:hintinfer}
%\end{figure}

@@

core.typed uses tools.analyzer to obtain a rich abstract syntax tree that includes resolved
reflection, however to tell the full story we include a rewrite relation

$$
\rewrite {\taenv{}} {\e{}} {\ep{}}
$$

to eliminate reflection when sufficient type hints are provided.

To help infer type hints

$$
\tajudgement {\taenv{}} {\hastype {\e{}} {\tatype{}}}
$$

gives the type hint \tatype{} of expression \e{} in environemtn \taenv{}.

$$
\TALetHint{}
$$

@@

  $$
    \RFieldElimRefl{}
  $$

@@

{\footnotesize
\input{constant}

\begin{figure*}
  $$
\begin{array}{lllll}
\updatefigure
\end{array}
$$
\caption{Type Update}
\end{figure*}
}

@@

\begin{figure}
  $$
\begin{array}{lllll}
  \restrictremovefigure{}
\end{array}
  $$
  \caption{Restrict and Remove}
\end{figure}
